---
title: "Customer satisfaction report"
author: "Emily Nordmann"
output: 
  word_document:
mainfont: Times
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,     # whether to show code chunks
  message    = FALSE,     # whether to show messages from your code
  warning    = FALSE,     # whether to show warnings from your code
  fig.width  = 8,         # figure width in inches (at 96 dpi)
  fig.height = 5         # figure height in inches (at 96 dpi)
)

library(tidyverse) 
library(patchwork) 
library(ggthemes)  
library(lubridate) 
library(knitr)
library(kableExtra)

survey_data <- read_csv("https://raw.githubusercontent.com/PsyTeachR/ads-v1/master/book/data/survey_data.csv")
```

# Overview

This report presents the results of the annual customer satisfaction survey at Goggles Inc.


* `caller_id` is a unique ID for each caller
* `employee_id` is a unique ID for each employee taking calls
* `call_start` is the date and time that the call arrived
* `wait_time` is the number of seconds the caller had to wait
* `call_time` is the number of seconds the call lasted after the employee picked up
* `satisfaction` is the customer satisfaction rating on a scale from 1 (very unsatisfied) to 5 (very satisfied)

## Call time versus wait time

This plot shows the relationship between how long a caller waits for their call to be picked up (shown on the x-axis) and how long their call lasted once the person taking the call picked up (show on the y-axis). The plot shows that there is a positive relationship between wait time and call time, that is, the longer people have to wait, the longer they then tend to spend on the call.  

```{r fig.cap = "Relationship between length of wait and subsequent call time"}
ggplot(data = survey_data, mapping = aes(x = wait_time, y = call_time)) +
  geom_point() + 
  geom_smooth(method = lm, 
              formula = y~x) +
  scale_x_continuous(name = "Wait Time (seconds)", 
                     breaks = seq(from = 0, to = 600, by = 60)) +
  scale_y_continuous(name = "Call time (seconds)",
                     breaks = seq(from = 0, to = 600, by = 30)) +
  theme_gdocs()
```

