# Conventions

This book will use the following conventions:

* Generic code: `list(number = 1, letter = "A")`
* Highlighted code: `r hl(list(number = 1, letter = "A"))`
* File paths: `r path("data/sales.csv")`
* R Packages: `r pkg("tidyverse")`
* Functions: `r hl(paste())`
* Strings: `r hl("Lisa")`, `r hl("Emily")`
* Numbers: `r hl(100)`, `r hl(3.14)`
* Logical values: `r hl(TRUE)`, `r hl(FALSE)`
* Glossary items: `r glossary("ordinal")`
* Citations: @R-tidyverse
* Internal links: Chapter\ \@ref(viz)
* External links: [Mastering Shiny](https://mastering-shiny.org/){target="_blank"}
* Menu/interface options: **`New File...`**
* Quiz question: I am going to learn a lot: `r torf(TRUE)`

```{r, echo=FALSE, results='asis'}
opts <- c(
   "the probability that the null hypothesis is true",
   answer = "the probability of the observed, or more extreme, data, under the assumption that the null-hypothesis is true",
   "the probability of making an error in your conclusion"
)

cat("* What is a p-value?", longmcq(opts))
```


```{block, webex.hide="Hidden Solutions"}
You found it!
```

::: {.info data-latex=""}
Informational asides.
:::

::: {.warning data-latex=""}
Notes to warn you about something.
:::

::: {.dangerous data-latex=""}
Notes about things that could cause serious errors.
:::

::: {.try data-latex=""}
Try it yourself.
:::

```{r}
# code chunks
paste("Applied", "Data", "Skills", 1, sep = " ")
```
