# Styling Plots {#plotstyle}

## Aesthetics

### Colour

The `colour` argument changes the point and line colour. Type `colours()` into the console to see a list of all the named colours in R. Alternatively, you can use hexadecimal colours like `"#FF8000"` or the `rgb()` function to set red, green, and blue values on a scale from 0 to 1.

Hover over a colour to see its R name.

<style>
  .colordemo { list-item-style: none; padding: 0; margin: 0;}
  
  .colordemo li, main ul.colordemo > li:first-child { 
    display: inline-block; 
    width: 0em; height: 0em; 
    border: 0.25px solid black; 
    margin: 0px; padding: 0.75em; 
    vertical-align: text-top;
  }
  .colordemo li span { 
    display: none; 
    background-color: white; 
    border: 0.25px solid black; 
    padding: 0.25em 0.5em; 
    border-radius: 0 1em 1em 1em;
  }
  .colordemo li:hover span { display: inline; position: absolute;}
</style>

```{r colour-demo, echo = FALSE, results='asis'}

name <- colours(TRUE)
rgb <- col2rgb(name) 
hsv <- rgb2hsv(rgb)
hex <- as.hexmode(rgb) %>% as.character() %>% 
  apply(2, paste, collapse = "")

colors <- tibble(name = name,
       hue = hsv[1, ],
       sat = hsv[2, ],
       val = hsv[3, ],
       hex = hex) %>%
  arrange(hue, sat, val)

glue::glue("    <li style='background-color: #{colors$hex};'><span>{colors$name}</span></li>") %>%
  paste(collapse = "\n") %>%
  paste0("\n<ul class='colordemo'>\n", ., "\n</ul>\n") %>%
  cat()
```

### Alpha

The `alpha` argument changes transparency (0 = totally transparent, 1 = totally opaque). 

```{r alpha-demo, echo = FALSE, fig.height = 3, fig.width = 10, fig.cap="Varying alpha values."}
alphas <- crossing(alpha = seq(0, 1, .1), y = 0:1)
ggplot(alphas, aes(x = alpha, y = y, alpha = I(alpha), 
                   color = as.factor(y))) +
  geom_hline(yintercept = 0, size = 5, color = "dodgerblue") +
  geom_hline(yintercept = 1, size = 5, color = "black") +
  geom_point(size = 20, show.legend = FALSE) +
  scale_x_continuous("alpha", breaks = alphas$alpha) +
  scale_y_continuous("", breaks = -1:2, limits = c(-1, 2)) +
  scale_color_manual(values = c("black", "dodgerblue")) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```

### Shape

The `shape` argument changes the shape of points. 

```{r shape-demo, echo = FALSE, fig.height = 2, fig.width = 10, fig.cap="The 25 shape values"}
shapes <- tibble(shape = 1:25, y = 0)
ggplot(shapes, aes(x = shape, y = y, shape = I(shape))) +
  geom_point(size = 8) +
  scale_x_continuous("", breaks = shapes$shape) +
  scale_y_continuous("", breaks = 0) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```

### Linetype

You can probably guess what the `linetype` argument does.


```{r linetype-demo, echo = FALSE, fig.height = 4, fig.width = 10, fig.cap="The 6 linetype values at different sizes."}
lines <- crossing(linetype = 1:6, size = c(0.25, 0.5, 1, 2)) %>% 
  mutate(x = linetype)
ggplot(lines) +
  geom_vline(aes(xintercept = linetype, 
                 linetype = I(linetype),
                 size = I(size))) +
  scale_x_continuous("linetype", breaks = 1:6, 
                     expand = expansion(add = 1)) +
  facet_wrap(~size, nrow = 1, labeller = label_both) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```
